---
title: 'P&S-2022: Lab assignment 2'
author: "Name1, Name2, Name3"
output:
  html_document:
    df_print: paged
---

```{r}
set.seed(18)

```

## General comments and instructions

-   Complete solution will give you $\bf 4$ points (out of 100 total). Submission deadline is **23:59 of 06 November 2022**\
-   The report must be prepared as an *R notebook*; you must submit to **cms** both the source *R notebook* **and** the generated html file\
-   At the beginning of the notebook, provide a work-breakdown structure estimating efforts of each team member\
-   For each task, include
    -   problem formulation and discussion (what is a reasonable answer to discuss);\
    -   the corresponding $\mathbf{R}$ code with comments (usually it is just a couple of lines long);\
    -   the statistics obtained (like sample mean or anything else you use to complete the task) as well as histograms etc to illustrate your findings;\
    -   justification of your solution (e.g. refer to the corresponding theorems from probability theory);\
    -   conclusions (e.g. how reliable your answer is, does it agree with common sense expectations etc)\
-   The **team id number** referred to in tasks is the **two-digit** ordinal number of your team on the list. Include the line **set.seed(team id number)** at the beginning of your code to make your calculations reproducible. Also observe that the answers **do** depend on this number!\
-   Take into account that not complying with these instructions may result in point deduction regardless of whether or not your implementation is correct.

### Task 1

#### In this task, we discuss the $[7,4]$ Hamming code and investigate its reliability. That coding system can correct single errors in the transmission of $4$-bit messages and proceeds as follows:

-   given a message $\mathbf{m} = (a_1 a_2 a_3 a_4)$, we first encode it to a $7$-bit *codeword* $\mathbf{c} = \mathbf{m}G = (x_1 x_2 x_3 x_4 x_5 x_6 x_7)$, where $G$ is a $4\times 7$ *generator* matrix\
-   the codeword $\mathbf{c}$ is transmitted, and $\mathbf{r}$ is the received message\
-   $\mathbf{r}$ is checked for errors by calculating the *syndrome vector* $\mathbf{z} := \mathbf{r} H$, for a $7 \times 3$ *parity-check* matrix $H$\
-   if a single error has occurred in $\mathbf{r}$, then the binary $\mathbf{z} = (z_1 z_2 z_3)$ identifies the wrong bit no. $z_1 + 2 z_2 + 4z_3$; thus $(0 0 0)$ shows there was no error (or more than one), while $(1 1 0 )$ means the third bit (or more than one) got corrupted\
-   if the error was identified, then we flip the corresponding bit in $\mathbf{r}$ to get the corrected $\mathbf{r}^* = (r_1 r_2 r_3 r_4 r_5 r_6 r_7)$;\
-   the decoded message is then $\mathbf{m}^*:= (r_3r_5r_6r_7)$.

#### The **generator** matrix $G$ and the **parity-check** matrix $H$ are given by

$$  
    G := 
    \begin{pmatrix}
        1 & 1 & 1 & 0 & 0 & 0 & 0 \\
        1 & 0 & 0 & 1 & 1 & 0 & 0 \\
        0 & 1 & 0 & 1 & 0 & 1 & 0 \\
        1 & 1 & 0 & 1 & 0 & 0 & 1 \\
    \end{pmatrix},
 \qquad 
    H^\top := \begin{pmatrix}
        1 & 0 & 1 & 0 & 1 & 0 & 1 \\
        0 & 1 & 1 & 0 & 0 & 1 & 1 \\
        0 & 0 & 0 & 1 & 1 & 1 & 1
    \end{pmatrix}
$$

#### Assume that each bit in the transmission $\mathbf{c} \mapsto \mathbf{r}$ gets corrupted independently of the others with probability $p = \mathtt{id}/100$, where $\mathtt{id}$ is your team number. Your task is the following one.

1.  Simulate the encoding-transmission-decoding process $N$ times and find the estimate $\hat p$ of the probability $p^*$ of correct transmission of a single message $\mathbf{m}$. Comment why, for large $N$, $\hat p$ is expected to be close to $p^*$.\
2.  By estimating the standard deviation of the corresponding indicator of success by the standard error of your sample and using the CLT, predict the \emph{confidence} interval $(p^*-\varepsilon, p^* + \varepsilon)$, in which the estimate $\hat p$ falls with probability at least $0.95$.\
3.  What choice of $N$ guarantees that $\varepsilon \le 0.03$?\
4.  Draw the histogram of the number $k = 0,1,2,3,4$ of errors while transmitting a $4$-digit binary message. Do you think it is one of the known distributions?

#### You can (but do not have to) use the chunks we prepared for you

#### First, we set the **id** of the team and define the probability $p$ and the generator and parity-check matrices $G$ and $H$

```{r}
# your team id number 
                          ###
id <- 50                  ### Change to the correct id!
                          ###
set.seed(id)
p <- id/100
# matrices G and H
G <- matrix(c(1, 1, 1, 0, 0, 0, 0,
		1, 0, 0, 1, 1, 0, 0,
		0, 1, 0, 1, 0, 1, 0,
		1, 1, 0, 1, 0, 0, 1), nrow = 4, byrow = TRUE)
H <- t(matrix(c(1, 0, 1, 0, 1, 0, 1,
		0, 1, 1, 0, 0, 1, 1,
		0, 0, 0, 1, 1, 1, 1), nrow = 3, byrow = TRUE))
# cat("The matrix G is: \n") 
#G  
#cat("The matrix H is: \n") 
#H
#cat("The product GH must be zero: \n")
#(G%*%H) %%2
```

#### Next, generate the messages

```{r}
# generate N messages

message_generator <- function(N) {
  matrix(sample(c(0,1), 4*N, replace = TRUE), nrow = N)
}  
messages <- message_generator(100)
codewords <- (messages %*% G) %% 2
```

#### Generate random errors; do not forget that they occur with probability $p$! Next, generate the received messages

```{r}
errors <- function(N){
  matrix(sample(c(0,1), 4*N, replace = TRUE), nrow = N)}
#  received <-   
#  
```

The next steps include detecting the errors in the received messages, correcting them, and then decoding the obtained messages. After this, you can continue with calculating all the quantities of interest

**Do not forget to include several sentences summarizing your work and the conclusions you have made!**

### Task 2.

#### In this task, we discuss a real-life process that is well modelled by a Poisson distribution. As you remember, a Poisson random variable describes occurrences of rare events, i.e., counts the number of successes in a large number of independent random experiments. One of the typical examples is the **radioactive decay** process.

#### Consider a sample of radioactive element of mass $m$, which has a big *half-life period* $T$; it is vitally important to know the probability that during a one second period, the number of nuclei decays will not exceed some critical level $k$. This probability can easily be estimated using the fact that, given the *activity* ${\lambda}$ of the element (i.e., the probability that exactly one nucleus decays in one second) and the number $N$ of atoms in the sample, the random number of decays within a second is well modelled by Poisson distribution with parameter $\mu:=N\lambda$. Next, for the sample of mass $m$, the number of atoms is $N = \frac{m}{M} N_A$, where $N_A = 6 \times 10^{23}$ is the Avogadro constant, and $M$ is the molar (atomic) mass of the element. The activity of the element, $\lambda$, is $\log(2)/T$, where $T$ is measured in seconds.

#### Assume that a medical laboratory receives $n$ samples of radioactive element ${{}^{137}}\mathtt{Cs}$ (used in radiotherapy) with half-life period $T = 30.1$ years and mass $m = \mathtt{team\, id \,number} \times 10^{-6}$ g each. Denote by $X_1,X_2,\dots,X_n$ the **i.i.d. r.v.**'s counting the number of decays in sample $i$ in one second.

1.  Specify the parameter of the Poisson distribution of $X_i$ (you'll need the atomic mass of *Cesium-137*)\
2.  Show that the distribution of the sample means of $X_1,\dots,X_n$ gets very close to a normal one as $n$ becomes large and identify that normal distribution. To this end,
    -   simulate the realization $x_1,x_2,\dots,x_n$ of the $X_i$ and calculate the sample mean $s=\overline{\mathbf{x}}$;
    -   repeat this $K$ times to get the sample $\mathbf{s}=(s_1,\dots,s_K)$ of means and form the empirical cumulative distribution function $\hat F_{\mathbf{s}}$ of $\mathbf{s}$;
    -   identify $\mu$ and $\sigma^2$ such that the \textbf{c.d.f.} $F$ of $\mathscr{N}(\mu,\sigma^2)$ is close to the \textbf{e.c.d.f.} $\hat F_{\mathbf{s}}$ and plot both **c.d.f.**'s on one graph to visualize their proximity (use the proper scales!);
    -   calculate the maximal difference between the two \textbf{c.d.f.}'s;
    -   consider cases $n = 5$, $n = 10$, $n=50$ and comment on the results.\
3.  Calculate the largest possible value of $n$, for which the total number of decays in one second is less than $8 \times 10^8$ with probability at least $0.95$. To this end,
    -   obtain the theoretical bound on $n$ using Markov inequality, Chernoff bound and Central Limit Theorem, and compare the results;\
    -   simulate the realization $x_1,x_2,\dots,x_n$ of the $X_i$ and calculate the sum $s=x_1 + \cdots +x_n$;
    -   repeat this $K$ times to get the sample $\mathbf{s}=(s_1,\dots,s_K)$ of sums;
    -   calculate the number of elements of the sample which are less than critical value ($8 \times 10^8$) and calculate the empirical probability; comment whether it is close to the desired level $0.95$

#### 1.Specify the parameter of the Poisson distribution:

$\nu = \lambda * N = \frac{ln(2)}{T[s]} * \frac{m}{M}N_A * teamId \approx 57564607 \approx 5.76*10^7$

```{r}
teamid = 18
lambda <- log(2)/(30.1*31536000)  # change this!
N <- teamid*6*10**17/137     # change this!
mu <- N * lambda
print(mu)
```

#### Simulate the realization of sample means

```{r}
set.seed(teamid)
K <- 1e3
n <- 50  # change N
sample_means <- colMeans(matrix(rpois(n*K, lambda = mu), nrow=n))
```

#### Next, calculate the parameters of the standard normal approximation

$$S_n =X_1 + X_2 + ... + X_n$$ $$M_n = S_n / n$$ As long as S is a sum of n i.i.d.r.v. X of Poisson distribution with paramater\
$\lambda$. $$\mu = E(X_i) = \lambda \ \ \ \ and  \ \ \ \sigma^2 =Var(X_i) = \lambda$$\
$$E(S_n) = E(\sum X_i) = n*E(X_i) = n\lambda$$\
$$\sigma^2 = Var(S_n) = Var(\sum X_i) = nVar(X_i) = n\lambda $$\
$$E(M_n) = E(S_n)/n = \lambda$$\
$$\sigma^2 = Var(M_n) = Var(S_n)/n^2 = \lambda/n$$\
Lets consider $Z_n = \frac{M_n - \lambda}{\sqrt{\frac{\lambda}{n}}} = \frac{M_n - \lambda}{\sqrt{\frac{\lambda}{n}}} \sim N(0,1)$\
By CLT we will get that M_n converges in low to Normal $$M_n \sim N(\lambda, \frac{\lambda}{n})$$\
And parameters are: $\mu = \lambda$ $\sigma=\sqrt{lambda/n}$\

```{r}


mu <- mu   # mu = n * mu / n
sigma <- sqrt(mu/n)   # sd = sqrt(Var[mean])/n = sqrt(n*Var[one]) / n = sqrt(mu/n)
```

#### We can now plot ecdf and cdf

```{r}
xlims <- c(mu-3*sigma, mu+3*sigma)
Fs <- ecdf(sample_means)

plot(Fs, 
     xlim = xlims, 
     ylim = c(0,1),
     col = "blue",
     lwd = 2,
     main = "Comparison of ecdf and cdf")
curve(pnorm(x, mean = mu, sd = sigma), col = "red", lwd = 2, add = TRUE)

```

#### The maximal difference between the two c.d.f.'s

```{r}
xl <- seq(mu-3*sigma, mu+3*sigma, by=0.01)

max_diff <- max(abs(ecdf(sample_means)(xl)-pnorm(xl, mean = mu,  sd = sigma)))

cat("For N =", n, "maximal difference between the two c.d.f.’s =", max_diff)
```

 #### Depending on different n

```{r}
mean_results_on_N <- function(n, K){
  teamid = 18
  set.seed(teamid)
  
  lambda <- log(2)/(30.1*31536000)
  N <- teamid*6*10**17/137
  mu <- N * lambda
  
  sample_means <- colMeans(matrix(rpois(n*K, lambda = mu), nrow=n)) # create samples ( s1, s2, s.. sk)

  # obtain Normal distr parameters
  mu <- mu   # change this!
  sigma <- sqrt(mu/n)   # change this!
  
  # get ecdf
  Fs <- ecdf(sample_means)
  
  # plot results
  xlims <- c(mu-3*sigma, mu+3*sigma)
  plot(Fs, 
       xlim = xlims, 
       ylim = c(0,1),
       col = "blue",
       lwd = 2,
       main = "Comparison of ecdf and cdf")
  # plot normal dist cdf
  curve(pnorm(x, mean = mu, sd = sigma), col = "red", lwd = 2, add = TRUE)
  
  # calculate maximal difference
  xl <- seq(mu-3*sigma, mu+3*sigma, by=0.01)
  max_diff <- max(abs(ecdf(sample_means)(xl)-pnorm(xl, mean = mu, sd = sigma)))
  cat("For N =", n, "maximal difference between the two c.d.f.’s =", max_diff, "\n")
  
  return(sample_means)
}
samples <-mean_results_on_N(5, 1e3)
samples <-mean_results_on_N(10, 1e3)
samples <-mean_results_on_N(50, 1e3)
```

#### Subtask2

```{r}
sum_results_on_N <- function(n, K){
  teamid = 18
  set.seed(teamid)
  
  lambda <- log(2)/(30.1*31536000)
  N <- teamid*6*10**17/137
  mu <- N * lambda
  
  sample_sums <- colSums(matrix(rpois(n*K, lambda = mu), nrow=n))

  mu <- mu   # change this!
  sigma <- sqrt(mu/n)   # change this!
  thres = 8*10**8
  emprob <- sum(sample_sums<tres)/K
  return(emprob)

}
x = 12
while(TRUE){
  curr_result <-sum_results_on_N(x, 1e3)
  cat("For n =", x, "Empirical Probability is ", curr_result, "\n")
  if (curr_result < 0.95){
    cat("Maximal value of n:", x-1)
    break
  }
  x = x + 1
}

```

#### Theoretical results

We have that $S =X_1 + X_2 + ... + X_n$ must be less than $t = 8*10^8$\

$$ P(S < t) = 1- P(S \geq t) \geq 0.95  \implies P(S\geq t) \leq 0.05$$\
As long as S is a sum of n i.i.d.r.v. X of Poisson distribution with paramater\
$\lambda$. $$\mu = E(X_i) = \lambda \ \ \ \ and  \ \ \ \sigma^2 =Var(X_i) = \lambda$$\
$$E(S) = E(\sum X_i) = n*E(X_i) = n\lambda$$\
$$\sigma^2 = Var(S) = Var(\sum X_i) = nVar(X_i) = n\lambda $$\
\##### Markov inequality\
$$ P(S \geq t)  \leq \frac{E(S)}{t} = \frac{n\lambda}{t}$$\
We know that higer bound probability must be less than 0.05\
$$ \frac{n\lambda}{t} \leq 0.05 \implies n \leq  \frac{0,05t}{\lambda} \approx 0.7$$\
As we can see Markov gives quite not an accurate result. It states that for all n P(S\<t) will be greater than 0.95\

##### Chebyshev inequality

$$P(S \geq t) = P(S - n\mu \geq t-n\mu) \leq P(|S-n\mu|\geq t-n\mu)$$\
Using Chebyshev inequality:\
$$ P(|S-n\mu|\geq t-n\mu) \leq (\frac{\sigma^2}{t-n\mu})^2 = (\frac{n\lambda}{t-n\lambda})^2$$ $$ (\frac{n\lambda}{t-n\lambda})^2 \leq 0.05 \implies n\lambda * ( 1 +\sqrt{0.05}) \leq t\sqrt{0.05} \implies n \leq 1.93$$\

#### Central Limit Theorem

$$P(S\leq t) = P( \frac{S-n\mu}{\sigma\sqrt{n}} \leq \frac{t-n\mu}{\sigma\sqrt{n}}) = Ф( \frac{t-n\mu}{\sigma\sqrt{n}}) \geq 0.95 \implies  \frac{t-n\mu}{\sigma\sqrt{n}} \geq  Ф^{-1}(0.95) $$\
$Ф^{-1}(0.95) \approx 1.67$ and we got a quadratic inequality:\
$$ n\lambda+ 1.67*\sqrt{n\lambda} \leq t$$\
That gives the higher bound $n \approx 13.7 \approx 14$\

##### Summary

AS we can see the results given with Markov inequality (0) are very not accurate and do not give us any bounds. Chebyshev inequality gives better approximation (2) but still not very accurate. And the central Limit Theorem gives results very close to the experimental ones (Both are equal to 14)

#### Conclusion

Having done experiments with divisions of radioactive elements that have Poisson distribution with specific parameters, we have got an interesting results

Firstly, We got that our  theoretical approximation is really close to  the experimental one even for small number (n=5) of samples. And for bigger ones (from 30) the difference is very neglectable. 
  
For N = 50 maximal difference between the two c.d.f.’s = 0.02251375  with number of samples K = 1000

Also we compared theoretical approximations for probabilities of exceeding the limit number of divisions in 1 second.
  1. The maximal value safe of n is 14, that corresponds with CLT results and experimental ones.
    - Markov inequality gives very inaccurate results
    - Chebyshev inequality gives better approximation
    - Central Limit Theorem gives results very close to the experimental ones


### Task 3.

#### In this task, we use the Central Limit Theorem approximation for continuous random variables.

#### One of the devices to measure radioactivity level at a given location is the Geiger counter. When the radioactive level is almost constant, the time between two consecutive clicks of the Geiger counter is an exponentially distributed random variable with parameter $\nu_1 = \mathtt{team\,id\,number} + 10$. Denote by $X_k$ the random time between the $(k-1)^{\mathrm{st}}$ and $k^{\mathrm{th}}$ click of the counter.

1.  Show that the distribution of the sample means of $X_1, X_2,\dots,X_n$ gets very close to a normal one (which one?) as $n$ becomes large. To this end,
    -   simulate the realizations $x_1,x_2,\dots,x_n$ of the \textbf{r.v.} $X_i$ and calculate the sample mean $s=\overline{\mathbf{x}}$;\
    -   repeat this $K$ times to get the sample $\mathbf{s}=(s_1,\dots,s_K)$ of means and then the \emph{empirical cumulative distribution} function $F_{\mathbf{s}}$ of $\mathbf{s}$;\
    -   identify $\mu$ and $\sigma^2$ such that the \textbf{c.d.f.} of $\mathscr{N}(\mu,\sigma^2)$ is close to the \textbf{e.c.d.f.} $F_{\mathbf{s}}$ of and plot both \textbf{c.d.f.}'s on one graph to visualize their proximity;\
    -   calculate the maximal difference between the two \textbf{c.d.f.}'s;\
    -   consider cases $n = 5$, $n = 10$, $n=50$ and comment on the results.
2.  The place can be considered safe when the number of clicks in one minute does not exceed $100$. It is known that the parameter $\nu$ of the resulting exponential distribution is proportional to the number $N$ of the radioactive samples, i.e., $\nu = \nu_1*N$, where $\nu_1$ is the parameter for one sample. Determine the maximal number of radioactive samples that can be stored in that place so that, with probability $0.95$, the place is identified as safe. To do this,
    -   express the event of interest in terms of the \textbf{r.v.} $S:= X_1 + \cdots + X_{100}$;\
    -   obtain the theoretical bounds on $N$ using the Markov inequality, Chernoff bound and Central Limit Theorem and compare the results;\
    -   with the predicted $N$ and thus $\nu$, simulate the realization $x_1,x_2,\dots,x_{100}$ of the $X_i$ and of the sum $S = X_1 + \cdots + X_{100}$;\
    -   repeat this $K$ times to get the sample $\mathbf{s}=(s_1,\dots,s_K)$ of total times until the $100^{\mathrm{th}}$ click;\
    -   estimate the probability that the location is identified as safe and compare to the desired level $0.95$

#### First, generate samples an sample means:

```{r}
teamid = 18
nu1 <- 10  + teamid  # change this!
K <- 1e3
n <- 5
sample_means <- colMeans(matrix(rexp(n*K, rate = nu1), nrow=n))
```

#### Next, calculate the parameters of the standard normal approximation

$$S_n =X_1 + X_2 + ... + X_n$$\
$$M_n = S_n / n$$\

As long as S is a sum of n i.i.d.r.v. X of Exponential distribution with paramater\
$\lambda$. $$\mu = E(X_i) = 1/ \lambda \ \ \ \ and  \ \ \ \sigma^2 =Var(X_i) = 1/\lambda^2$$\
$$E(S_n) = E(\sum X_i) = n*E(X_i) = n/\lambda$$\
$$\sigma^2 = Var(S_n) = Var(\sum X_i) = nVar(X_i) = n/\lambda^2 $$\
$$E(M_n) = E(S_n)/n = 1/\lambda$$\
$$\sigma^2 = Var(M_n) = Var(S_n)/n^2 = 1/\lambda n$$\
$$ Z_n = \frac{M_n - \mu}{\sigma} = \frac{M_n - 1/\lambda}{\sqrt{1/\lambda n}}$$\
By CLT we will get that M_n converges in low to Normal $$M_n \sim N(\frac{1}{\lambda}, \frac{1}{n\lambda})$$\
And parameters are: $\mu = 1/\lambda$ $\sigma= 1/\sqrt{\lambda n}$\

```{r}
mu <- 1/nu1   # change this!
sigma <- 1/(nu1*sqrt(n))   # change this!
```

#### We can now plot ecdf and cdf

```{r}
xlims <- c(mu-3*sigma,mu+3*sigma)
Fs <- ecdf(sample_means)
plot(Fs, 
     xlim = xlims, 
     col = "blue",
     lwd = 2,
     main = "Comparison of ecdf and cdf")
curve(pnorm(x, mean = mu, sd = sigma), col = "red", lwd = 2, add = TRUE)


xl <- seq(mu-3*sigma, mu+3*sigma, by=0.01)
max_diff <- max(abs(ecdf(sample_means)(xl)-pnorm(xl, mean = mu, sd = sigma)))
cat("For N =", n, "maximal difference between the two c.d.f.’s =", max_diff, "\n")
```

```{r}
mean_results_on_N <- function(n, K){
  teamid = 18
  set.seed(teamid)
  nu1 <- teamid + 10
  
  sample_means <- colMeans(matrix(rexp(n*K, rate = nu1), nrow=n))

  # obtain Normal distr parameters
  mu <- 1/nu1   # change this!
  sigma <- 1/(nu1*sqrt(n))   # change this!
  
  # get ecdf
  Fs <- ecdf(sample_means)
  
  # plot results
  xlims <- c(mu-3*sigma, mu+3*sigma)
  plot(Fs, 
       xlim = xlims, 
       ylim = c(0,1),
       col = "blue",
       lwd = 2,
       main = "Comparison of ecdf and cdf")
  # plot normal dist cdf
  curve(pnorm(x, mean = mu, sd = sigma), col = "red", lwd = 2, add = TRUE)
  
  # calculate maximal difference
  xl <- seq(mu-3*sigma, mu+3*sigma, by=0.01)
  max_diff <- max(abs(ecdf(sample_means)(xl)-pnorm(xl, mean = mu, sd = sigma)))
  cat("For N =", n, "maximal difference between the two c.d.f.’s =", max_diff, "\n")
  
  return(sample_means)
}
res <- mean_results_on_N(5, 1e3)
res <- mean_results_on_N(10, 1e3)
res <-mean_results_on_N(50, 1e3)
```
#### Subtask2

```{r}
sum_results_on_N <- function(N, K){
  teamid = 18
  set.seed(teamid)
  n <- 100
  nu <- (teamid + 10)*N
  sample_sums <- colSums(matrix(rexp(n*K, rate = nu), nrow=n))

  thres <- 1
  eprob <- sum(sample_sums<thres)/K
  return (eprob)
}
print(sum_results_on_N(4, 1e3))

x = 6
while(TRUE){
  curr_result <-sum_results_on_N(x, 1e3)
  cat("For n =", x, "Empirical Probability is ", curr_result, "\n")
  if (curr_result < 0.95){
    cat("Maximal value of n:", x+1)
    break
  }
  x = x - 1
}

```
#### Theoretical part

We have that $S =X_1 + X_2 + ... + X_n$ must be less than $t = 1$\

$$ P(S < t) = 1- P(S \geq t) \geq 0.95  \implies P(S\geq t) \leq 0.05$$\
As long as S is a sum of n i.i.d.r.v. X of Poisson distribution with paramater\
$\lambda = 28N$. $$\mu = E(X_i) = 1/\lambda \ \ \ \ and  \ \ \ \sigma^2 =Var(X_i) = 1/\lambda^2$$\
$$E(S) = E(\sum X_i) = n*E(X_i) = n/\lambda$$\
$$\sigma^2 = Var(S) = Var(\sum X_i) = nVar(X_i) = n/\lambda^2 $$\
##### Markov inequality\
$$ P(S \geq t)  \leq \frac{E(S)}{t} = \frac{n}{t\lambda}$$\
Higher bound of probability must be < 0.05\
$$ \frac{n}{t\nu N} \leq 0.05 \implies N \geq  100*20t\nu \approx 56000$$\

##### Chebyshev inequality

$$P(S \geq t) = P(S - n\mu \geq t-n\mu) \leq P(|S-n\mu|\geq t-n\mu)$$\
Using Chebyshev inequality:\
$$ P(|S-n\mu|\geq t-n\mu) \leq (\frac{\sigma^2}{t-n\mu})^2 = (\frac{n/\lambda^2}{t-n/\lambda})^2$$ \
$$ (\frac{n/\lambda^2}{t-n/\lambda})^2 \leq 0.05 \implies (\frac{100/N^2\nu^2}{t-100/N\nu})^2 \leq 0.05 \implies ... \implies N \leq 1200$$\

#### Central Limit Theorem

$$P(S\leq t) = P( \frac{S-n\mu}{\sigma\sqrt{n}} \leq \frac{t-n\mu}{\sigma\sqrt{n}}) = Ф( \frac{t-n\mu}{\sigma\sqrt{n}}) \geq 0.95 \implies  \frac{t-n\mu}{\sigma\sqrt{n}} \geq  Ф^{-1}(0.95) $$\

$$\frac{t-n/\lambda}{\sqrt{n}/\lambda} \geq  Ф^{-1}(0.95)$$\
$Ф^{-1}(0.95) \approx 1.67$ 

$$ 100/N\nu+ 16.7/N\nu \leq 1$$\
That gives the higher bound $n \approx 4.83 \approx 5$\

#### Summary
As we can see the experimental results are very close to that ones that we obtained in theory even when the number of samples is low.
The distribution of times for random samples tends to the Normal Distribution with specified parameters and the error is minimal.
Also, we can see that approximations with Markov inequality is useless and do to gives any reasonable boundaries. Chebyshev gives better results but still not so accurate. However, the CLT calucation returns very accurate limit for `n`.




**Do not forget to include several sentences summarizing your work and the conclusions you have made!**

### General summary and conclusions

SUmmarize here what you've done, whether you solved the tasks, what difficulties you had etc

### General summary and conclusions

SUmmarize here what you've done, whether you solved the tasks, what difficulties you had etc

We've solved all three tasks. We've explored how random events affect some of the real-life examples.

Generally, theoretical work took much more time than coding.

The difficulties: chebushevs inequality, trying to write one-liners in 'R style', matrices
